---
title: "data_wrangling_i"
output: html_document
date: "2023-09-19"
---

---
title: "Data Wrangling"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Let's import `the FAS_litters.csv` csv using a relative path. 

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") #path name come from the "data" folder that we created in the repo folder on deskstop, that's why we put it there

litters_df = 
  janitor::clean_names(litters_df)   #replaces the df with updated version. we could load the library called "janitor", but we dont need allll the functions so we are just calling clean_names by using the :: notation
```

---
title: "Data Wranling document"
output: github_document
---

```{r}
library(tidyverse)
```

Let's import `the FAS_litters.csv` csv using a relative path. 

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") #path name come from the "data" folder that we created in the repo folder on deskstop, that's why we put it there

litters_df = 
  janitor::clean_names(litters_df)   #replaces the df with updated version. we could load the library called "janitor", but we dont need allll the functions so we are just calling clean_names by using the :: notation
```

Import the same dataset using an absolute path 
# watch lecture on this


look at data summary
```{r}
str(litters_df)
```

# Options in `read_*`

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv" , skip = 10 , col_names = FALSE) #makes up column names but...
```

Look at NA values

```{r}
litters_df = 
  read_csv(
    "data/FAS_litters.csv" ,
    na = c("NA", 19)) #anywhere that NA or 19 show up, it is treated like a missing value
```


Column types 

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv" , 
           col_types = 
             cols(
               #FIND CODE FOR THIS
             )) #list column and what it's supposed to be

```


## Other file types 

Import a clsc file first

```{r}
mlb_df = 
  read_excel("data/mlb11.xlsx")
```

Import SAS file 
```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")
```

## Base r

```{r, eval = FALSE}
litters_df = 
  read.csv("data/FAS_litters.csv")

litters_df$Gr #DONT do this
```

## Export data


```{r}
litters_df_cleaned = 
  read_csv("data/FAS_litters.csv")

litters_df_cleaned =
  janitor::clean_names(litters_df_cleaned)

write_csv(litters_df_cleaned, "data/litters_cleaned.csv")
```
